Data Management \\ and LSST Special Programs

### 2023-05-02 ap_procbounds.tex
Update ap_procbounds.tex

Note that processing boundaries might ultimately be defined not by what is technically possible, but by the resulting image quality parameters (e.g., the number of stars with sufficient flux for photometric calibration).
Furthermore, the processing boundaries might not be fully constrained until the final performance of the LSST Science Pipelines, as described in the Data Management Applications Design, \citeds{LDM-151}) document, is fully characterized.

{\bf Summary of the conclusions below:}\\
Very short ($<$2 sec) exposures could be difficult to reduce due to an incompletely-formed PSF, and very short or very long ($>$150 sec) exposures could be difficult to calibrate due to having too few (or too few unsaturated) stars.
It is currently unclear whether images with very bright sky backgrounds (twilight images) can be processed with the LSST Science Pipelines, or whether user generated pipelines will be needed.
The full reduction and calibration of images obtained with non-sidereal tracking, in which the stars are streaked, is currently beyond the scope of the LSST Science Pipelines, and will require a user generated pipeline.

@@

\subsubsection{Short Exposures (Non-Standard Visits of $<$30 sec)}
The LSST System Requirements document states that {\it ``The LSST shall be capable of obtaining and processing exposures not taken in a standard visit mode including those with a minimum exposure time of} {\tt minExpTime}", which is 1 second (stretch goal 0.1 seconds; LSR-REQ-0111 in \citeds{LSE-29}).

However, for exposure times there are other considerations, as changing the exposure time also affects the photometric and astrometric calibrations.
Assuming that 1 second exposure can be reduced and calibrated, its detected point sources will span a dynamic range of $r$$\approx$ 13--21 magnitudes.
A template image built on 15 second exposures will saturate at $r$$\approx$15.8 mag, but this still leaves stars between 15.8--21.0 magnitudes to be used in the PSF-matching (and all other filters have a similarly large overlap).

@@

% In conversation with DM-AP team members (Reiss, Findeisen, Connolly, Bo) there has not yet been a study of the safe range of exposure times that will be allowed to contribute Alert Production.
% One possibly useful study is Chang et al. (2012), "Atmospheric point spread function interpolation for weak lensing in short exposure imaging data".
% They show that a 15 second exposure contains PSF variability on short spatial scales across a 1 square degree image which, for extragalactic fields with few stars (i.e., but good for weak lensing), is hard to characterize.
% They also present a new software package to do mitigate the effects.
% Software packages \texttt{PhoSim} (Peterson et al. 2015; \citep{2015ApJS..218...14P}) or \texttt{ARROYO} \citep{2004SPIE.5497..290B} could be used to characterize the PSF stability as a function of exposure time.

@@

There is no maximum exposure time specified for an LSST image.
Given that the template image will be a stack of at least a year or two of data, processing a $5$--$10$ times deeper single image through the difference imaging pipeline should be fine.
However, a $2\times150$ second exposure would saturate at $r \approx 18.3$, perhaps leaving too few stars overlapping with e.g., templates or WFD images, for astrometric and photometric calibrations.
Furthermore, cosmic-ray rejection completeness might be reduced for longer exposures (unknown), which could impact the quality of a difference image and the detected sources.
Additionally, any system qualities that vary on short (but $>30$ second) timescales could inhibit photometric calibration (e.g., tracking).

@@

Images obtained during twilight for scientific purposes are also likely to have shorter exposure times, and so the issues described in Section \ref{ssec:procbounds_expt} also apply here.
Whether or not bright-background images can (or shall) be fully processed -- reduced, calibrated, background-subtracted, and delivered with astrometric and photometric solutions -- or whether this will require a user generated pipeline, remains to be determined (see also the example in Section \ref{ssec:SPCS_Twilight}).
This may depend on the exposure time and the number of stars available in the image.

@@

Non-sidereal tracking leads to images in which stars are streaked, but the moving object appears as a point source.
Full processing -- providing reduced, calibrated, background-subtracted images that are delivered with astrometric and photometric solutions -- of these images is beyond the scope of the DM pipelines as it would require the development of new algorithms, and will need to be done as a user generated pipeline.
The first steps of such a pipeline, such as Instrument Signature Removal, will probably be possible to achieve by reconfiguring the relevant DM software tasks.

%%% MLG removed the following in Feb 2022, it's no longer a concern.
% \subsection{Number of Exposures per Visit (Long Sequences of a Single Field)}

% There is no processing constraint on the number of consecutive exposures that could be obtained of a single field.
% From a DM perspective, it would be best if these exposures were packaged into visits of no more than 2 exposures per visit, to minimize the need to reconfigure of the pipelines, and because the camera only ``clears" between visits.

% K.-T. Lim has pointed out that an odd number of exposures is a non-standard visit; two snaps is hardwired into the code. This is baked-in to a configuration so that the pipeline can have a definition of what kind of timing delay constitutes ``late".  Moving away from 2 exposures per visit requires a configuration change to the pipelines, which incurs an overhead (up to 1 minute) -- in fact, K.-T. things that between $10$ and $120$ seconds exposure times can easily be handled by the pipeline (i.e., can be run through ISR using scaled calibration frames), so long as they come in pairs. The real problem is knowing how long the processing should take, and not killing a process that is taking longer because there were 4 snaps in the visit instead of 2. To accommodate non-standard visits requires that the scheduler pass on the information of the number of snaps in the visit (\ref{DMSR-1}). Then the processing pipeline will know to, e.g., not attempt to difference the two snaps in the case were there is an odd number of snaps in a visit. \textit{MLG -- I've heard rumors of a CR regarding alternate standard visits of $1\times30$ seconds, but do not know the status or implications of this.}

% K.-T. has also pointed out that currently, a deep drilling field would be interpreted as a single visit of 50 exposures by the scheduler. One implication of this is that since the camera only ``clears" prior to a new visit, it would not do this for the entire 50-exposure sequence. The processing pipeline would need to know how to divide this sequence up into visits. As there is no current requirement for DM to receive the information that the scheduler is about to do a 50 exposure visit, we need \ref{DMSR-1} to add the proposal ID and the number of exposures per visit to the meta data, and then it should be OK for DM to parse this visit information in the reduction pipeline.

%%% MLG removed the following in Feb 2022, it's no longer a concern for Special Programs (crowded fields are in the WFD).
% \subsection{Images in Very Crowded Fields}

% The LSST pipelines' performance in crowded fields is documented in \citeds{DMTN-077}, which finds that, e.g., in Galactic Plane regions with a source density of $500000$ sources per square degree, the completeness drops to 50\% at $20.2$ magnitudes.
% The slide deck at \citeds{Document-27962} also describes DM's plans for processing crowded fields. These may or may not be appropriate for Special Programs data, depending on the science goals.

>>>>

Note that processing boundaries might ultimately be defined not by what is technically 
possible, but by the resulting image quality parameters, e.g., the number of stars with 
sufficient flux for photometric calibration.

Furthermore, the processing boundaries might not be fully constrained until the final 
performance of the LSST Science Pipelines, as described in the Data Management 
Applications Design, \citeds{LDM-151}) document, is fully characterized.

{\bf Summary.}\\

\begin{itemize}
\item \bf{Short or long exposures.} Very short ($<$2 sec) exposures could be difficult to 
process due to an incompletely-formed PSF. Short (or long, $>$150 sec) exposures 
could be difficult to calibrate due to having too few (or too few unsaturated) stars.
\item \bf{Bright backgrounds (twilight).} It is currently unclear whether images with very 
bright sky backgrounds (twilight images) can be processed with the LSST Science Pipelines, 
or whether user generated pipelines will be needed.
\item \bf{Non-sidereal tracking.} The full reduction and calibration of images obtained 
with non-sidereal tracking, in which the stars are streaked, is currently beyond the scope 
of the LSST Science Pipelines, and will require a user generated pipeline.
\end{itemize}

@@

\subsubsection{Short Exposures (Non-Standard Visits of $<$30 sec)}

The LSST System Requirements document states that {\it ``The LSST shall be capable of 
obtaining and processing exposures not taken in a standard visit mode including those 
with a minimum exposure time of} {\tt minExpTime}", which is 1 second (stretch goal 0.1 
seconds; LSR-REQ-0111 in \citeds{LSE-29}).

However, for exposure times there are other considerations, as changing the exposure time 
also affects the photometric and astrometric calibrations.
Assuming that 1 second exposure can be reduced and calibrated, its detected point sources 
will span a dynamic range of $r$$\approx$ 13 to 21 magnitudes.
A template image built on 15 second exposures will saturate at $r$$\approx$15.8 mag, but 
this still leaves stars between 15.8 and 21.0 magnitudes to be used in the PSF-matching 
(and all other filters have a similarly large overlap).

@@

@@

There is no maximum exposure time specified for an LSST image.
Given that the template image will be a stack of at least a year or two of data, 
processing a $5$--$10$ times deeper single image through the difference imaging pipeline 
should be fine.

However, a $2\times150$ second exposure would saturate at $r \approx 18.3$, perhaps 
leaving too few stars overlapping with e.g., templates or WFD images, for astrometric and 
photometric calibrations.

Furthermore, cosmic-ray rejection completeness might be reduced for longer exposures 
(unknown), which could impact the quality of a difference image and the detected sources. 
Additionally, any system qualities that vary on short (but $>30$ second) timescales could 
inhibit photometric calibration (e.g., tracking).

@@

Images obtained during twilight for scientific purposes are also likely to have shorter 
exposure times, and so the issues described in Section \ref{ssec:procbounds_expt} also 
apply here.

Whether or not bright-background images can (or shall) be fully processed -- reduced, 
calibrated, background-subtracted, and delivered with astrometric and photometric 
solutions -- or whether this will require a user generated pipeline, remains to be 
determined (see also the example in Section \ref{ssec:SPCS_Twilight}). 
This may depend on the exposure time and the number of stars available in the image.

@@

Non-sidereal tracking leads to images in which stars are streaked, but the moving object 
appears as a point source.

Full processing -- providing reduced, calibrated, background-subtracted images that are 
delivered with astrometric and photometric solutions -- of these images is beyond the 
scope of the DM pipelines as it would require the development of new algorithms, and will 
need to be done as a user generated pipeline. 
The first steps of such a pipeline, such as Instrument Signature Removal, will probably 
be possible to achieve by reconfiguring the relevant DM software tasks.
